---
title: "MachineLearning"
author: "Sandra Müller"
date: "1 2 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task from course website

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

## Data loading
```{r}
df <- read.csv("pml-training.csv")
colnames(df)
#summary(df)
str(df)
```

## Setup

```{r}
library(caret)
set.seed(42)
train <- createDataPartition(y=df$classe,p=.70,list=F)
training <- df[train,]
testing <- df[-train,]
```
## Cleanup

Model fitting requires less than 53 variables, so cleaning up the data for variables which don't contribute or only contribute little 
```{r}

#exclude columns with lot of NAs
nas <- colSums(is.na(training))/nrow(training)

nACols <- which(nas > 0.75)
cleanTraining <-  training[,-nACols]

#remove administrative columns and timestamps
cleanTraining <- cleanTraining[,-grep("X|name|cvtd_timestamp|raw_timestamp_part_1|raw_timestamp_part_2|classe|new_window|num_window",names(cleanTraining))]

#remove vars with little variance
cleanTrainingF <- cleanTraining[,-nearZeroVar(cleanTraining)]

#str(cleanTraining)
length(names(cleanTrainingF))
```

# Fitting model
```{r}
library(randomForest)
modelFit <- randomForest(training$classe ~ ., data=cleanTrainingF)
print(modelFit)
```
## Cross validation & expected out of sample error
```{r}
library(caret)
pred <- predict(modelFit, testing)
table(pred, testing$classe)
pRes <- postResample(pred, testing$classe)
pRes
confusionMatrix(pred, testing$classe)
```

## Apply to verifying dataset (don't put results in final html)
```{r message = FALSE}
verify <-  read.csv("pml-testing.csv",na.strings=c("NA","","#DIV/0!"))
pred <- predict(modelFit, newdata = verify, type="class")
#print(pred)
```
